On the Synchronization Power of the Openflow Data Plane

### 2015-03-22 inband-STN.tex
duplicated section 4.

\newcommand{\stefan}[1]{\textit{\textcolor{red}{[stefan]: #1}}} % Comments
\newcommand{\liron}[1]{\textit{\textcolor{green}{[julius]: #1}}} % Comments
\newcommand{\petr}[1]{\textit{\textcolor{blue}{[carlos]: #1}}} % Comments

@@

@@

\item by performing $\textit{unclaim}(i)$, $i\in[N]$, controller $i$
  \emph{unclaim} identifier $k$;

>>>>

\newcommand{\stefan}[1]{\textit{\textcolor{red}{[stefan]: #1}}} % Comments
\newcommand{\liron}[1]{\textit{\textcolor{green}{[liron]: #1}}} % Comments
\newcommand{\petr}[1]{\textit{\textcolor{blue}{[carlos]: #1}}} % Comments

@@

%\section{The Container Object}\label{sec:vision}
\section{-NEW- Transactional Synchronization}\label{sec:main}

We assume transactional access to  two kinds of objects: a
\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
given (bounded or unbounded) set, and a read-write \emph{memory} abstraction
object, which can be used for storing the current used policy identifier.
%
\subsubsection{A resource claimer}
A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
\emph{unclaim} and \emph{abortIfClaimed} with the following sequential
semantics:
%The container synchronization object introduced in this paper
%is essentially a transactional memory abstraction which can be implemented
%in standard Openflow switches.
%Before presenting the technical details in the next section,%
%we give a brief overview.

\item by performing $\textit{unclaim}(k)$, $i\in[N]$, $k\in\Nat$, controller $i$
  \emph{unclaim} identifier $k$;

\item $\textit{abortIfClaimed}(k)$, $k\in\Nat$, 
abort current transaction if $k$ is currently claimed by any controller.
%returns $\textit{true}$ if and
%  only if $k$ is not currently claimed by any controller; if
%  $\textit{false}$ is returned, we say that the check operation
%  \emph{fails} (which causes an abort of the transaction invoking it).

\subsubsection{A memory abstraction}
A memory can be accessed with the following sequential
semantics:

\begin{itemize}
\item by performing $\textit{Write}(addr,k)$,
  $addr,k\in\Nat$, the value at memory offset $addr$ is set to $k$;

\item $\textit{abortIfUnequal}(addr,k)$, $addr,k\in\Nat$, 
aborts current transaction if the value at memory offset $addr$ is not equal to $k$.
%returns $\textit{true}$ if and
%  only if $k$ is not currently claimed by any controller; if
%  $\textit{false}$ is returned, we say that the check operation
%  \emph{fails} (which causes an abort of the transaction invoking it).

Note that reading the memory can be done by just reading the switch configuration however such action doesn't affect the execution of the transaction (as $\textit{abortIfUnequal}$ does).

An immediate extension of the memory interface is the \emph{compare-and-set} (CAS) operation that takes two integer parameters
$\textit{old}$ and $\textit{new}$, reads the register and, if the
value of it is $\textit{old}$, replaces it with $\textit{new}$ and
returns $\textit{true}$; otherwise, $\textit{false}$ is returned and
we say that the CAS operation \emph{fails} (which causes an abort if
the transaction invoking it).

\liron{this subsection doesn't flow well for me}

We now describe how to implement our transactional 
abstractions using standard OpenFlow features.

%features , including
%\emph{bundling}, \emph{conditional application}, and \emph{groups}.
%[[PK let us leave for the moment
%We will describe later how to implement some of these features
%in older Openflow versions where they are not
%supported. \textbf{TODO}
%]]
Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of  
transactional operations  $\textit{claim}(x)$, $\textit{check}(x)$, $\textit{unclaim}(x)$,
and $\textit{CAS}(x)$, respectively. 
Algorithm~\ref{alg:commit} describes the procedure executed by
a transaction at \emph{commit time}.

We describe each operation as a message factory that returns OpenFlow commands that should be sent to the switch as part of the same transaction.
We assume that every transaction has a unique identifier  used to start a \emph{bundle}. 
All messages generated during a transaction invocation, are wrapped by bundle messages carrying the bundle's id and followed by bundle commit message.
Messages with same bundle id are considered to belong to same bundle and are performed atomically.

Our implementation is based on OpenFlow flow entries modification command, the FlowMod message that explained in Section \ref{sec:background}. We use the CHECK\_OVERLAP optional flag to make the transaction execution dependent on switch state.

We consider the match part of an entry to be a ternary string of bounded length (represented by binary value and mask) which can be applied to any set of supported packet header fields. In our implementation the $64$bit metedata field alone is suffice. Moreover we consider the action part of an entry to be an integer, which in practice can be implemented by a Set-Metadata instruction where the written value is that integer.

  \liron{got till here}

We use creation and deletion of flow entries as abstraction to claiming/unclaiming  resource id numbers and also to checking whether a given id is claimed. The match part of the added flow entry represents the added
number and the mask part of the match varies from one controller to another in order to allow multiple controllers to claim the same number without overriding the same flow entry. 
As can be seen in Algorithms \ref{alg:claim}-\ref{alg:unclaim}, the match mask is a concatenation of the controller id with all ones, while the value is the id concatenated twice. We can see that for any two numbers $x_1$ and $x_2$ and controller ids $c_1\neq c_2$ the match patterns $m_1=(x_1\concat x_1, c_1\concat all-1)$ and $ m_1=(x_2\concat x_2, c_2\concat all-1)$ are not equal even if $x_1=x_2$ therefore they can both claim and unclaim without affecting each other.

@@

\item by performing $\textit{unclaim}(k)$, $i\in[N]$, $k\in\Nat$, controller $i$
  \emph{unclaim} identifier $k$;
